Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-10-29 manuscript.tex
latest

In ideal environments, we predict evolution for higher $b$ and $c$, but not lower $d$. In plants and corals, the ability to secure territories $c$ is strongly associated with rapid individual growth; not surprisingly, rapid individual growth is a defining feature of the competitor trait cluster \cite{grime_1977,darling_2012}. The analogue of Grime's ideal environment for fishes occurs when resources limitations are not severe and the environment varies predictably over large temporal and spatial scales \cite{winemiller_1992,winemiller_2015}. The associated ``periodic'' strategy is characterized by enormous brood sizes, suggesting a strategy of ensuring that more propagules actually end up contesting areas favorable for development (higher $b$). Evolution for higher $b$ (which is closely related to the intrinsic low-density growth rate $r=b-d$) will be discussed further in the Discussion in the context of MacArthur's $r$/$K$ scheme.

@@

>>>>

In ideal environments, we predict evolution for higher $b$ and $c$, but not lower $d$. In plants and corals, a key mechanism for winning territorial contests is rapidly outgrowing and ``shading out'' competitors; not surprisingly, rapid individual growth is a defining feature of the competitor trait cluster \citep{grime_1977,darling_2012}. Evolution for higher $b$ under high-density, competitive conditions may seem counter-intuitive given that $b$ is closely related to the low-density reproductive rate $r=b-d$, and the opposite outcome is expected in MacArthur's $r$/$K$ dichotomy \citep{macarthur_1962}. Neither particularly high nor low $b$ have been associated with the competitor strategy in plants and corals. However, for fishes, the analogous ``periodic'' strategy is characterized by enormous brood sizes as well as rapid development \citep{winemiller_1992,winemiller_2015}, suggesting a strategy of ensuring that many propagules actually end up contesting areas favorable for development (higher $b$). The evolution of $b$ in ideal environments will be discussed further in the Discussion.

@@

The model developed here is a density-dependent generalization of the classic lottery model (Results section).
